An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2014-05-12 paper.tex
added GraphBuilder and GraphLab refs, beefed up the ACG section.

\subsection{ACG}
\label{sec:acg}

What people popularly consider Big Data applications fall into two broad
categories. First, data processing tasks that fit into the MapReduce model
where data is tagged and sorted to discover relationships. These sorts of
applications only require scale out rather than scale up. Scaling out requires
replicas to process data simultaneously, but do not need to coordinate for that
data processing. Scaling up, what scientific simulations do, requires sometimes
serious coordination between the data processing processes for any of them to
succeed. The second category is more similar to scientific simulations in that
it requires information from other processes to succeed. These graph-style
applications can be done using Hadoop, but it does not fit as well.

To address these graph applications, GraphLab and GraphBuilder are evaluated
to determine how to extend the IO stack to better support these applications.
GraphBuilder is a set of MapReduce tasks that extract, normalize, partition and
serialize a graph out of unstructured data, and writes graph-specific formats
into HDFS. These files are later consumed by GraphLab, a vertex-centric,
asynchronous execution engine that runs directly on top of HDFS (i.e.,
non-MapReduce). The following illustrates the architecture of both frameworks:

@@

In order to make both work on top of the exascale stack, both have to be
modified. After these modifications are implemented, GraphBuilder will
be able to write the partitioned graph in the newly proposed HDF5
files which will thus be stored in the IOD nodes (or IONs) in a
parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
the library to perform at high speeds by benefiting from the new
features, such as the function shipping. In general both frameworks will be
modified so that calls to HDFS-based formats are replaced by the proposed HDF5
ones. This is referred to as the HDF Adaptation Layer or HAL and will
provide, from the GraphBuilder/GraphLab point of view:

@@

The initial phase of this project has determined the necessary changes in the
HDF5 format to support these features. These identified features will be
proven during phase 2 with a demonstration of GraphBuilder and GraphLab.

@@

All of the tests are performed on the Buffy Cray test cluster at Los Alamos
National Laboratory. It consists of ......

>>>>

\subsection{Arbitrarily Connected Graphs}
\label{sec:acg}

What people popularly consider Big Data applications fall into two broad
categories. First, data processing tasks that can fit into the MapReduce model
where data is tagged and sorted to discover relationships. These sorts of
applications only require scale out rather than scale up. Scaling out requires
replicas to process data simultaneously, but do not need to coordinate for that
data processing. Scaling up, what scientific simulations do, requires sometimes
serious coordination between the processes for any of them to succeed. In the
middle are graph applications that, with some replicated data, can be made to
fit reasonably well into the MapReduce model. The challenge is having access to
the edge and vertex lists effectively to build partitions for independent
processing. GraphBuilder~\cite{Jain:2013:GraphBuilder} is a tool to generate
effective graph paritions reducing the load for using MapReduce to process
graph data sets. GraphLab~\cite{Low:2012:GraphLab} offers a way to process
these graphs efficiently for parallel platforms using a minimum of
communication. Using these tools as motivators, changes to the HDF5 interface
and the underlying storage infrastructure is proposed.  The following
illustrates the architecture of both frameworks:

@@

In order to make both of these tools work on top of the exascale stack, they
both have to be modified. After these modifications are implemented,
GraphBuilder will be able to write the partitioned graph in the newly proposed
HDF5 files which will thus be stored in the IOD nodes (or IONs) in a
parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
the library to perform at high speeds by benefiting from the new
features, such as the function shipping. In general both frameworks will be
modified so that calls to HDFS-based formats are replaced by the proposed HDF5
equivalents. This is referred to as the HDF Adaptation Layer or HAL and will
provide, from the GraphBuilder/GraphLab point of view:

@@

The initial phase of this project has determined the necessary changes in the
HDF5 format to support these features. These identified features will be
proven during phase two with a demonstration of GraphBuilder and GraphLab.

@@

All of the tests are performed on the Buffy Cray test cluster at Los Alamos
National Laboratory. It consists of 84 nodes each with dual, 8 core Intel
Xeon ES-2670 CPUs at 2.6 GHz and 32 GB RAM. The inteconnect is Cray Gemini.
The storage array are three sets of SSD-based lustre file systems.
